---
title: "Project one"
author: "Eric Anku"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import-libraries}
# Import libraries for data analysis

pacman::p_load(
        tidyverse, #Data management and plots
        rio, # Data import
        here, #Project management
        janitor, #Clean data names
        skimr #Skim dataset
)
```

```{r load-dataset}
df <- import(here("data","birthweight_dataset.sav")) #Import dataset
```



```{r clean-data}
df <- clean_names(df) #clean data variables
```

```{r glimpse-data}
glimpse(df)  #Glimpse data frame
```

```{r data-transformation}
df <- df%>%mutate(
        smoker = factor(smoker)%>%
                fct_recode("No smoker" = "0",
                           "Smoker" = "1"),
        lowbwt = factor(lowbwt)%>%
                fct_recode("No" = "0",
                           "Yes" = "1"),
        mage35 = factor(mage35)%>%
                fct_recode("No" = "0",
                           "Yes" = "1")
)
```

```{r skim-data}
df_no_id <- df%>%select(-c(id))
skim(df_no_id)
```


```{r subset-data-smoking-mothers}
df_smoker <- df%>%
        filter (smoker == "Smoker") ## Data frame for smoking mothers
```


```{r subset-data-non-smoking-mothers}
df_non_smoker <- df%>%
        filter(smoker == "No smoker")
```



#### 1. What is the mean birthweight of babies of non-smoking mothers?

The mean birth weight of babies of non-smoking mothers is `r round(mean(df_non_smoker$birthweight),2)` kg.


#### 2. What is the mean birthweight of babies of smoking mothers?

The mean birth weight of babies of non-smoking mothers is `r round(mean(df_smoker$birthweight),2)` kg.


#### 3. What is the mean head circumference for babies of non-smoking mothers?

The mean head circumference of babies of non-smoking mothers is `r mean(df_non_smoker$headcirc)` cm.

## 4. What is the mean gestational age at birth for babies of smoking mothers?


The mean gestational age of babies of smoking mothers is `r mean(df_smoker$gestation)` weeks.


## 5. What is the maximum head circumference for babies of smoking mothers?

The maximum head circumference 

```{r}
max(df_smoker$headcirc)
```


## 6. What is the minimum gestational age at birth for babies of smoking mothers?

## 7. What is the baby birth weight range for babies of smoking mothers?

## 8. Are head circumference data for babies of smoking mothers normally distributed?

## 9. What is significance value for the above Shapiro-Wilk test?

## 10. How are birthweight data of non-smoking mothers skewed?

## 11. Are birthweight data for babies of smoking mothers normally distributed?


```{r}

```

